---
title: "r-101-grade"
author: "Iris Mancini"
format: html
editor: visual
code-fold: true
---

```{r, message = FALSE}
here::i_am("r-101-grade-analysis.Rproj")
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(vroom)
library(stringr)
```

### Question 1 - Data loading :

```{r, message = FALSE}

grades <- vroom(here("grades.csv"))
```

### Question 2 - 

```{r, message = FALSE, echo = FALSE}
summarize <- grades |> 
  summarise(Minimum = min(Exam, na.rm=TRUE), Maximum = max(Exam, na.rm=TRUE), Median = median(Exam, na.rm=TRUE), Mean = mean(Exam, na.rm=TRUE))

summarize |>
  knitr::kable()
  
```

### Question 3 - 

```{r, message = FALSE, echo = FALSE}
exam_na <-  grades |>
  filter(is.na(Exam)) |>
  count(is.na(Exam))
```

There are `r exam_na` students who didn't take the Exam.

### Question 4 -

```{r, message = FALSE}
grades |> 
  drop_na(Exam) |>
ggplot(aes(x = Exam)) + 
  geom_bar(fill = "lightblue", color = "white") + 
  xlab("Grade at the Exam") + ylab("number of students") + 
  labs(title = "Distribution of the grades") + 
   theme(plot.title = element_text(hjust = 0.5)) 
```

We can see that the distribution is more concentrated around lower values. There aren't a lot of students who got very good grades.

### Question 5 - 

```{r, message = FALSE}
group_df <- grades |>
  group_by(Group) |>
  summarise(N=n()) 
group_df |>
  knitr::kable()
```

### Question 6 - 

```{r, message = FALSE}

#je crée une variable IndexGroup, qui prendra les valeurs des numéros de groupe pour rendre le graphique plus lisible
grades$IndexGroup <- as.numeric(gsub("grp_", "", grades$Group))

grades |>
  ggplot(aes(x=factor(IndexGroup), fill = factor(IndexGroup))) + 
  geom_bar() + 
  labs(title = "Number of students in each group") + 
  theme(plot.title = element_text(hjust = 0.5),  legend.position="none") + 
  ylab("number of students") + 
  xlab("group ID")

  

```

The number of students per group is fairly evenly distributed

### Question 7 - 

1st way of representing the distribution :

```{r, message = FALSE, warning = FALSE}

#on attribue un numéro à chaque groupe (son index)
grades2 <- grades |>
  mutate(IndexGroup = as.numeric(gsub("grp_", "", Group)))

#on réordonne la variable Group selon la variable IndexGroup
grades2 <- grades |>
  mutate(Group = factor(Group, levels = unique(Group[order(IndexGroup)])))

ggplot(grades2, aes(x = Exam, y=Group, fill = Group)) +
  geom_boxplot() + theme(legend.position="none") + 
  ggtitle("distribution of exam scores by group") + 
  theme(plot.title = element_text(hjust = 0.5))
```

The group with the best exam average is group 21.

2.  An other way of plotting the distribution :

```{r, message = FALSE, warning = FALSE}
ggplot(grades, aes(x = Exam, fill = Group)) + 
  geom_histogram(bins = 15) + 
  geom_rug(alpha = 0.5, linewidth = 1.5) + 
  theme(legend.position="none", strip.text = element_text(size = 10)) + 
  ggtitle("distribution of exam scores by group") + 
  facet_wrap(~Group)
```

After eliminating the legend, which displays informations we already have access to with the title of each graph in the facet_wrap, the whole graphical representation is more readible. We find again that students in the group 21 (which we saw had the better mean in the precedent graph), have grades that are more located around higher values of grades. It helps us understand better the information displayed in the boxplot.

### Question 8 - 

```{r, message = FALSE}
n_group <- grades |>
  group_by(Group) |>
  summarise(exam_na = sum(is.na(Exam)))

knitr::kable(n_group)

```

### Question 9 - 

```{r}

#je créer une variable qui me permet d'associer un numéro à chaque groupe, afin de faciliter la lisibilité du graphique ensuite
n_group$indexgroup2 <- as.numeric(gsub("grp_", "", n_group$Group))

ggplot(n_group, aes(x= factor(indexgroup2), y = exam_na, fill = Group, color = "orange")) + 
  geom_col() + 
  ggtitle("Number of students who missed the Exam in each group") +
  ylab("number of students") + 
  xlab("Group number") +
  theme(plot.title  = element_text(size = rel(0.9)), legend.position="none", strip.text = element_text(size = 10)) 
```

### Question 10 - 

```{r, message = FALSE}

grades_long_format <- grades |>
  pivot_longer(
   cols = -c(Id, Group), #les deux seules variables que l'on ne va pas transformer
   names_to = "name", 
   values_to = "value"
  )
```

### Question 11 - 

```{r}

id_exam_na <-  grades_long_format |>
  group_by(Id)|>
  summarise(NbExamsMissed = sum(is.na(value))) 
```

### Question 12 - 

```{r, message = FALSE}

ggplot(id_exam_na, aes(x =NbExamsMissed)) + 
  geom_histogram(binwidth = 1, fill = "skyblue", color = "black") + 
  ggtitle("distribution of missing values of grades") +
  ylab("number of students") + 
  xlab("total number of missing values") + 
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))
  
```

### Question 13 - 

```{r, message = FALSE}

exam_na2 <- grades_long_format |> 
  filter(name=="Exam") |>
  group_by(Group) |>
  summarise(NbOfStudentsMissing = sum(is.na(value))) 
  
exam_na2 |>
  knitr::kable()

```

### Question 14 - 

```{r, message = FALSE}


online_na <- grades_long_format |>
  filter(str_starts(name, "Online_MCQ")) |>
  group_by(Id) |>
  summarise(NbOfNa_online = sum(is.na(value)))

  
```

### Question 15 -

```{r, message = FALSE, warning = FALSE}

grades_na_online <- grades |>
  rowwise() |>
  mutate(NbOnlineMissed = sum(is.na(c_across(starts_with("Online_MCQ")))) )

grades_na_online|> 
  ggplot(aes(x = Exam)) +
  geom_bar(fill = "skyblue", color = "darkblue") + 
  ylab("missing grades at the online test") +
  xlab("Exam score") + 
  ggtitle("exam score conditioned by the number of online missing grades") +
  theme(plot.title = element_text(hjust = 0.5))
```

We can clearly conclude that there is a strong correlation between missing online exams and having a bad grade at the exam. We don't know if there's a causal effect however.

### Question 16 - 

```{r, message = FALSE}

grades_na_online2 <- grades |>
  rowwise() |>
    mutate(OneOnlineNA = any(is.na(c_across(starts_with("Online_MCQ"))))) |>
  select(Id, OneOnlineNA)
```

### Question 17 - 

```{r, message = FALSE}

grades_na_online3 <- grades |>
  rowwise() |>
    mutate(OneOnlineNA = any(is.na(c_across(starts_with("Online_MCQ")))))

percentage_tab <- grades_na_online3 |>
  group_by(Group) |>
  mutate(nb_group = n()) |>
  mutate(P_missed = sum(OneOnlineNA)/nb_group*100) |>
  select(Group, P_missed)
```

### Question 18 - 

```{r, message = FALSE}

mean_exam <- grades |>
  group_by(Group) |>
  summarise(average = mean(Exam, na.rm = TRUE)) 

```

on voit que les élèves sont un peu nuls

```{r, message = FALSE}

whole_table <-  inner_join(percentage_tab, mean_exam, by="Group")
```

Représentation graphique :

```{r, message = FALSE}

ggplot(whole_table, aes(x=average, y = P_missed)) +  geom_jitter(alpha = 0.03, width = 0.01, height = 0.01) + 
  ggtitle("Colleration between average exam score of each group and the % of students who missed at least one online test") + 
  ylab("% of students who missed at least one online test")+   xlab("average score at the exam of the group") + 
   theme(axis.title.y = element_text(size = 8), plot.title = element_text(size = rel(0.7))) 

    
  
```

There is relationship between the groups who have a higher value of the share of students who missed at least one online test and the average score at the exam of the group.
